### 2013-11-30 paper/main.tex
added abstract, minor cleanups

% {{{ abstract
%\pagebreak
%\thispagestyle{empty}
%\begin{abstract}
%\noindent

% TOOD

% What methods are used? PID, direct? ...

%\end{abstract}
% }}}

@@

The engine model used here is based work
by Butts and Sivashankar\autocite{532315} which is derived from
the work by Powell and Cook\autocite{4789342}.
The engine configuration is a modern 4.6L V-8.
To simplify analysis the linearized model is used as shown in
Figure \ref{fig:lem}.

@@

Then substituting them in to Equation \ref{eq:dd} results in
the simplified from of Equation \ref{eq:lems}.

\begin{align}
	\frac{y}{u} &= \frac{(1.699) (8.5683) (2.98) (1)}
		{(z^2 - 0.9025z)(z - 0.9354)(1) + (8.5683)(2.98)(0.00093)} \notag \\
	\Aboxed{ \frac{y}{u} &= \frac{43.38}{z^3 - 1.838 z^2 + 0.8442 z + 0.02375}}
		\label{eq:lems}
\end{align}

\begin{samepage}
It is still necessary to address the delta inputs and outputs.
It can be seen in Figure \ref{fig:nodelta} there is a beneficial
side effect that makes this trivial.  The transform that converts
from steady state to a delta cancels
with the transform that converts from a delta to steady state
\footnote{If the torque input was non zero the transforms would not
cancel.}.  Therefore the final transform is still Equation \ref{eq:lems}.
An example of the output response with no control is shown in
Figure \ref{fig:em_plot}.
\end{samepage}

@@

\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/engine_model_plot}
\end{center}
\caption{Output response of engine model with no control.
The control can range from zero to one.}
Here the control is constant at 0.65 which resulted in a stable
rpm near 900.
The Matlab script is provided in Appendix \ref{app:em_matlab}.
\label{fig:em_plot}
\end{figure}

% see octave/engine_model.m for these calculations

>>>>

% {{{ abstract
%\pagebreak
%\thispagestyle{empty}
\begin{abstract}
\noindent
The task of maintaining a stable idle for an internal combustion
engine with spark ignition is non-trivial.
Any time an accessory is turned on/off the torque applied to
the engine changes.
And changes in torque will change the engine rpm if the control
inputs are constant.
This paper shows how control systems methods can be applied to
the problem of idle stabilization.
Methods include: pole placement, PID, direct design, and
various state space designs.

\end{abstract}
% }}}

@@

The engine model used here is based work
by Butts and Sivashankar\autocite{532315} which was derived from
the work by Powell and Cook\autocite{4789342}.
The configuration is a modern 4.6L V-8 gas engine.
A single control input representing the idle air bypass valve

To simplify analysis the linearized model is used as shown in
Figure \ref{fig:lem}.

@@

\begin{samepage}
Then substituting them in to Equation \ref{eq:dd} results in
the simplified from of Equation \ref{eq:lems}.
\begin{align}
	\frac{y}{u} &= \frac{(1.699) (8.5683) (2.98) (1)}
		{(z^2 - 0.9025z)(z - 0.9354)(1) + (8.5683)(2.98)(0.00093)} \notag \\
	\Aboxed{ \frac{y}{u} &= \frac{43.38}{z^3 - 1.838 z^2 + 0.8442 z + 0.02375}}
		\label{eq:lems}
\end{align}
\end{samepage}

\begin{samepage}
It is still necessary to address the delta inputs and outputs.
It can be seen in Figure \ref{fig:nodelta} that there is a beneficial
side effect which makes this trivial.
The transform that converts from steady state to a delta cancels
with the transform that converts from a delta to steady state
\footnote{In the more general case, when torque is non zero,
the transforms would not cancel.}.
Therefore the final transform is still Equation \ref{eq:lems}.
An example of the output response with no control is shown in
Figure \ref{fig:em_plot}.
\end{samepage}

@@

\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/engine_model_plot}
\end{center}
\caption{Output response of engine model with no control.
The control can range from zero to one.
Here the control is constant at 0.65 which resulted in a stable
rpm near 900.\protect\footnotemark}
\label{fig:em_plot}
\end{figure}
\footnotetext{The Matlab source to produced these plots is in Appendix \ref{app:em_matlab}.}
